### 2024-05-27 thm1.2.tex
thm1.2

Theorem~\ref{thm:main-recur} implies the existence of a recursive algorithm $\A$ for $k$-SJ. For 2-SJ, we can set $\A$ to the ``folklore algorithm'' mentioned in Section~\ref{sec:intro:prev}, which ensures $F_2(n, \out) = O(n \log n + \out)$. For $k \geq 3$, $\A$ recursively solves $O(k^3)$ subproblems of $k-1$-SJ---each of which has an input size of at most $n$ and an output size of at most $\out$---and then produces the output for $k$-SJ using $O(n\log n + k\cdot \out)$ additional processing
time. It is rudimentary to prove that $\A$ is comparison-based by induction.
Moreover, we have the following recurrence.
For $k \geq 3$, there is a constant $c$ such that
\myeqn{
    F_{k}(n, \out) = c \cdot k^3 \cdot \big( F_{k-1}(n, \out) + n \log n + k \cdot \out \big).\nn
}
We solve the recurrence by expanding it $k-2$ times. And we get
\myeqn{
    F_{k}(n,\out) &=& c \cdot k^3(n\log n + k\cdot \out )+ \nn \\
    && c^2 \cdot k^3(k-1)^3 \cdot (F_{k-2}(n,\out) + n\log n + k\cdot \out)\nn\\
    &\leq& (c+c^2)\cdot k^3(k-1)^3(n\log n + k \cdot\out)+  \nn\\
    && c^2 \cdot k^3(k-1)^3 \cdot F_{k-2}(n,\out) \nn \\
    &\leq& (c+c^2+c^3)\cdot k^3(k-1)^3(k-2)^3(n\log n + k \cdot\out)+  \nn\\
    && c^3 \cdot k^3(k-1)^3(k-2)^3 \cdot F_{k-3}(n,\out)\nn\\
    &\leq& ... \nn \\
    &\leq& (\sum_{i =1}^{k-2}c^i)\cdot (k!/2)^3\cdot (n\log n + k \cdot\out)+ \nn \\
    && c^{k-2}(k!/2)^3 \cdot F_{2}(n,\out) \nn \\
    &\leq & 2c^{k-1}(k!/2)^3(n\log n + k \cdot\out) \nn \\
    &=& O(c^k \cdot (k!)^3(n\log n + k \cdot\out)). \nn
}
As a result, Theorem~\ref{thm:main-alg} holds.