### 2024-05-28 appB.tex
yufei

\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:recur-output}.}

\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:properties}.}

\extraspacing {\bf Computing $\bm{R^*_1, R^*_2, ..., R^*_{k-2}}$.}

\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:cross-r*}.}

\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type2:recur-output}.}

\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type2:properties}.}

\extraspacing {\bf Computing $\bm{H^*}$.}

\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type2:contained-h*}.}