### 2024-05-28 appB.tex

\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:recur-output}.}
 We will prove by associating each tuple $(r_1',...,r_{k-2}')\in \J(R_1',...,R_{k-2}')$ with a unique tuple in the H-V $k$-SJ result set $\J(R_1,...,R_{k-2},H,V)$. More specifically, let $r_i$ be the full rectangle of $r_i'$ for $i\in[k-2]$. Let $r'_l = \gleft(\bm{t'})$ and $r'_b = \gbot(\bm{t'})$, where $l,b\in[k-2]$. Let $h$ be the segment $h_\bot$ for the rectangle $r_b$, and $v$ be the segment $v_\vdash$ for the rectangle $r_l$. The intersection of $h$ and $v$ is the bottom-left coner of the intersection area $\bigcap_{i = 1}^{k-2}r_i'$. Hence, $h\cap v\cap \bigcap_{i = 1}^{k-2}r_i \neq \emptyset$ and the tuple $(r_1,...,r_{k-2},h,v)$ is a result tuple in $\J(R_1,...,R_{k-2},H,V)$. The result tuple is unique because for any two distinct tuples $\bm{t'_1}$ and $\bm{t'_2}$ in $\J(R_1',...,R_{k-2}')$, there exists an integer $i \in [k-2]$ s.t. the full rectangles of $\bm{t_1'}[i]$ and $\bm{t_2'}[i]$ are different. Therefore, $|\J(R_1',...,R_{k-2}')| \leq |\J(R_1,...,R_{k-2})| = \out$.

\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:properties}.}

\extraspacing {\bf Computing $\bm{R^*_1, R^*_2, ..., R^*_{k-2}}$.}

\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:cross-r*}.}

\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type2:recur-output}.}

\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type2:properties}.}

\extraspacing {\bf Computing $\bm{H^*}$.}

\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type2:contained-h*}.}