### 2024-05-28 4-comp.tex
4-comp

We will first compute $\maxtop(r')$ for each rectangle $r' \in \bigcup_{i = 1}^{k-2}R_{i}'$ by scanning the set $\J(R_1',...,R_{k-2}')$. Initially, we set $\maxtop(r') = -\infty$ for each $r' \in \bigcup_{i = 1}^{k-2}R_{i}'$. For each tuple $\bm{t} = (r_1',...,r_{k-2}') \in \J(R_1',...,R_{k-2}')$, we update $\maxtop(\gbot(\bm{t}))$ to the maximum between its current value and $\ytop(B_{\bm{t}})$. Equipped with the values of $\maxtop(r')$ for each $r' \in \bigcup_{i = 1}^{k-2}R_{i}'$, we can construct the set $R_i^*$ for each $i \in [k-2]$ by collecting the top-sliced rectangles for all rectangles in $R_i'$.