### 2024-05-30 4.3.tex
yufei

\noindent \underline{\em Proof of Statement (2).} 
=======
\vgap

\noindent \underline{\em Proof of Statement (2).}

% Next, we prove (2).  Since $\gbot(\bm{t}) = r'$, $\dcross_H(\bm{t})$ contains exactly the set of segments segment $h \in H$ that cross the rectangle $[x_1,x_2]\times [y_1, \ytop(B_\bm{t})]$, where $\ytop(B_{\bm{t}})\leq \maxtop(r')$. Therefore, $\dcross_H(\bm{t})$ $\subseteq \cross_H(r^*)$. Furthermore, if the (horizontal) segments of $\cross_H(r^*)$ are sorted in ascending order of y-coordinate, then $\dcross_H(\bm{t})$ includes a prefix of the sorted order.

Let $r' = [x_1,x_2]\times [y_1,y_2]$, and $r^* = [x_1,x_2] \times [y_1,\maxtop(r')]$. We will prove that $\dcross_H(\bm{t})$ contains exactly the set of segments $h\in H$ that cross the rectangle $[x_1,x_2]\times [y_1,\ytop(B_\bm{t})]$.
>>>>>>> 4b540a0e220d934a24e26c2f624815c59364ca4c