### 2024-05-30 4.3.tex
yufei

>>>>>>> fc9fcecf3343c248a654ce6af7c44ce99af1e489
\noindent \underline{\em Proof of Statement (2).}