### 2024-05-30 4.3.tex
4.3

\noindent \underline{\em Proof of Statement (2).}