### 2024-05-30 4.3.tex

<<<<<<< HEAD

=======
>>>>>>> 792e1cb443a6d771c5dbf687d0745b6bf4017a9a
\noindent \underline{\em Proof of Statement (2).}