### 2024-05-30 4.3.tex
4.3

\noindent \underline{\em Proof of Statement (2).} 
Let $r' = [x_1,x_2] \times [y_1,y_2]$. Then, $r^* = [x_1,x_2]\times [y_1, \maxtop(r')]$. Let $h$ be an arbitrary segment in $\dcross_H(\bm{t})$, whose y-coordinate is $y_h$.  We will prove that $h$ crosses $r^*$, which leads to $\dcross_H(\bm{t})\subseteq \cross_H(r^*)$. It suffies to show that $\xleft(h) < x_1 \le x_2 < \xright(h)$ and $y_h \in [y_1, \maxtop(r')]$. 
\begin{itemize}
    \item Since $h$ crosses $r' = [x_1,x_2]\times [y_1,y_2]$, we have $\xleft(h) < x_1 \le x_2 < \xright(h)$.
    \item Since $h$ crosses $B_{\bm{t}}$, we have $y_h \in [y_1, \ytop(B_{\bm{t}})]$. By the definition of $\maxtop(r')$ (see Definition \eqref{eqn:hv:type1:maxtop}), and the fact that $r' = \gbot(B_{\bm{t}})$, we have $\ytop(B_{\bm{t}}) \leq \maxtop(r')$. Therefore, $y_h \in [y_1, \maxtop(r')]$.
\end{itemize}

\vgap

Then, we will prove that each segment $h \in \cross_H(r^*)$ satisfying that the y-coordinate $h$ is no more than $\ytop(B_{\bm{t}})$ is in $\dcross_H(\bm{t})$. This claim leads to the last sentence of Statement (3). Consider an arbitrary $h\in \cross_H(r^*)$, whose y-coordinate $y_h$ satisfies that $y_h \le \ytop(B_{\bm{t}})$. We will prove that $h$ crosses $r'$ and $B_{\bm{t}}$, and therefore $h \in \dcross_H(\bm{t})$.

\vgap

First, we prove that $h$ crosses $r'$. It suffices to show that $\xleft(h)  < x_1 \leq x_2 < \xright(h)$ and $y_h \in [y_1, y_2]$. Both conditions are satisfied since $h$ crosses $r^* = [x_1,x_2]\times [y_1, \maxtop(r')]$, and $\maxtop(r')\leq y_2$. Then, we prove that $h$ crosses $B_{\bm{t}}$. It suffices to show that $\xleft(h) < \xleft(B_{\bm{t}}) \le \xright(B_{\bm{t}}) < \xright(h)$ and $y_h \in [y_1, \ytop(B_{\bm{t}})]$.
\begin{itemize}
    \item Since $h$ crosses $r'$, we have $\xleft(h) < x_1 \leq x_2 < \xright(h)$. Since $r' = \gbot(B_{\bm{t}})$, we have the x-range of $B_{\bm{t}}$ is contained in $[x_1,x_2]$. Therefore, $\xleft(h) < \xleft(B_{\bm{t}}) \le \xright(B_{\bm{t}}) < \xright(h)$.
    \item Since $h$ crosses $r^*$, we have $y_h \in [y_1, \maxtop(r')]$. By the definition of $h$, we have $y_h \leq \ytop(B_{\bm{t}})$. Therefore, $y_h \in [y_1, \ytop(B_{\bm{t}})]$. 
\end{itemize}