### 2024-05-31 5.2.tex
5.2

\noindent \underline{\em Proof of Statement (1).} Consider any tuple $\bm{t} = (r_1,...,r_{k-3}, h', v) \in \J(R_1,...,R_{k-3},H',V)$. Let $h$ be the full segment of $h'$. Since $\bm{t} \in \J(R_1,...,R_{k-3},H',V)$, $B_\bm{t}$, which is the intersection point of $h'$ and $v$, is in $r_i$ for $i \in [k-3]$. As $h$ is the full segment of $h'$, $B_\bm{t} \in h$. Since $v$ also contains $B_\bm{t}$, $B_\bm{t}$ is also the intersection of $h$ and $v$. Let $r_{k-2}$ be an arbitrary rectangle in $\contained_{R_{k-2}}(\bm{t})$. By the definition of $\contained_{R_{k-2}}(\bm{t})$, $r_{k-2}$ covers effective segment of $h'$, whose left enpoint is $B_{\bm{t}}$. Therefore, $B_{\bm{t}}$ is in $r_{i}$ for $i \in [k-2]$, and $\bm{t} \in \J(R_1,...,R_{k-2},H,V)$. Moreover, since the left endpoint of the effective segment of $h'$ is the same as the left endpoint of $h$, $r_{k-2}$ contains the left endpoint of $h$. Therefore, $\bm{t} \in \J_2$.

\vgap

\noindent \underline{\em Proof of Statement (2).}

\vgap

\noindent \underline{\em Proof of Statement (3).}

\vgap